### 2014-02-19 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Reworking intro paragraphs in dpp manuscript.

\section*{Introduction}
Understanding the processes that generate biodiversity and regulate
community assembly is a major goal of evolutionary biology.
Given the dynamic nature of our planet, a major component of this goal is
understanding how geological and climatic processes can affect the
evolutionary history of whole groups of co-distributed species and their
associated microbiota, including pathogens.
Such large-scale processes predict patterns of shared evolutionary history
across affected populations of organisms, e.g., temporally clustered
divergences across co-distributed species.
At recent timescales, such events should leave a signature in the genetic
variation within and among the affected lineages.
Thus, accurately estimating models of shared evolutionary history across
co-distributed taxa is important for better understanding how regional and
global biogeographic processes affect biodiversity.

@@

The software package \msb implements such an approach in an
approximate-Bayesian model-choice framework \citep{Hickerson2006,Huang2011}.
Given aligned sequence data for \npairs{} pairs of populations, \msb 
estimates the number of divergence events shared among the pairs, the timing of
the events, and the assignment of pairs to the events, while integrating out
uncertainty in demographic parameters and the genealogical history of the
sequences.
% samples over all the possible ways of assigning the divergences between the
% populations of each pair to 1--\npairs{} divergence events, while estimating
% the timing of the events and integrating out uncertainty in demographic
% parameters and the genealogical history of the sequences.
Thus, the method samples over all possible divergence models of differing
dimensionality (i.e., all the possible partitions of \npairs{} pairs to
$1,2,\ldots,\npairs{}$ divergence-time parameters), and, in so doing, estimates
the posterior probability of each model.

@@

We modify the \msb model to test whether alternative parameterizations and
priors will improve the behavior of the approximate-Bayesian model-choice
approach.
We implement a Dirichlet-process prior (DPP) over all possible models of
divergence, and use gamma and beta probability distributions in place of
uniform priors on many of the model's parameters.
Using simulations, we show that the new implementation has improved robustness,
accuracy, and power compared to the original model.
Also, our results confirm that the improved performance of the new model
is due to a combination of
(1) more flexible priors on divergence-time and demographic parameters that
avoid placing high prior density in improbable regions of parameter space, and
(2) a diffuse Dirichlet-process prior that does not strongly disfavor
divergence models with intermediate numbers of divergence events.
% Also, our results confirm that both the uniform priors on demographic and
% divergence-time parameters and the U-shaped prior on divergence models
% contribute to the tendency of \msb.
After reanalyzing sequence data from 22 pairs of taxa from the Philippines
\citep{Oaks2012} under the new model, we find a large amount of posterior
uncertainty in the number of divergence events shared among the taxa; a result
in contrast with the original \msb model \citep{Oaks2012} and congruent with
intuition given the richness of the model and the relatively small amount of
information in the data.

>>>>

\section*{Introduction}

Understanding the processes that generate biodiversity and regulate
community assembly is a major goal of evolutionary biology.
Large-scale changes to the environment, including geological and climatic
events, can affect the evolutionary history of entire communities of
co-distributed species and their associated microbiota.
For example, by partitioning communities, such an event can isolate groups of
populations and cause a temporal cluster of speciation events across
co-distributed taxa.
Given the dynamic nature of our planet, such biogeographical processes likely
play a significant role in determining diversification rates and patterns.
At recent timescales, temporal clusters of diversification caused by
biogeographical events can leave a signature in the genetic variation within
and among the affected lineages.
Thus, methods for accurately estimating models of shared evolutionary events
across co-distributed taxa from genetic data are important for better
understanding how regional and global biogeographical processes affect
biodiversity.

@@

The software package \msb implements such an approach in an
approximate-Bayesian model-choice framework \citep{Hickerson2006,Huang2011}.
The method models temporally clustered divergences across taxa caused by a
biogeographical event (or a ``divergence event'') as a single, instantaneous
occurrence.
In other words, a divergence event causes a set of taxa to share the same
moment of divergence along a continuous time scale (i.e., simultaneous
divergence).
Given aligned sequence data for \npairs{} pairs of populations, \msb 
estimates the number of divergence events shared among the pairs, the timing of
the events, and the assignment of pairs to the events, while integrating out
uncertainty in demographic parameters and the genealogical histories of the
sequences.
% samples over all the possible ways of assigning the divergences between the
% populations of each pair to 1--\npairs{} divergence events, while estimating
% the timing of the events and integrating out uncertainty in demographic
% parameters and the genealogical history of the sequences.
Thus, the method samples over all possible divergence models of differing
dimensionality (i.e., all the possible partitions of \npairs{} pairs to
$1,2,\ldots,\npairs{}$ divergence events), and, in so doing, estimates
the posterior probability of each model.

@@

We modify the \msb model to test whether alternative parameterizations and
priors will improve the behavior of the approximate-Bayesian model-choice
approach.
We implement a Dirichlet-process prior (DPP) over all possible models of
divergence, and use gamma and beta probability distributions in place of
uniform priors on many of the model's parameters.
Using simulations, we show that the new implementation has improved robustness,
accuracy, and power compared to the original model.
Also, our results confirm that the improved performance of the new model
is due to a combination of
(1) more flexible priors on divergence-time and demographic parameters that
avoid placing high prior density in improbable regions of parameter space, and
(2) a diffuse Dirichlet-process prior that does not strongly disfavor
divergence models with intermediate numbers of divergence events.
% Also, our results confirm that both the uniform priors on demographic and
% divergence-time parameters and the U-shaped prior on divergence models
% contribute to the tendency of \msb.
After reanalyzing sequence data from 22 pairs of taxa from the Philippines
\citep{Oaks2012} under the new model, we find a large amount of posterior
uncertainty in the number of divergence events shared among the taxa; a result
in contrast with the original \msb model and congruent with
intuition given the richness of the model and the relatively small amount of
information in the data.
